---
title: "Assignment Statistical Inference - Part 1"
author: "OscarAMtz"
date: "8 de abril de 2019"
output:
  pdf_document: default
  geometry: margin=2cm
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Part 1
1. A simulation exercise
2. Basic inferential data analysis

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.
```{r}
set.seed(215)
lambda <- 0.2; obs <- 40; sims <- 1000 ## var simulations
exp.sims <- NULL ## var used to store exponential simulations
for (i in seq(sims)) exp.sims <- c(exp.sims, mean(rexp(obs, lambda)))
exp.sims.mean <- mean(exp.sims)
round(exp.sims.mean, 3)
```
The mean of the 1000 simulations is `r round(exp.sims.mean, 3)`, meanwhile the theorical mean is `r 1/lambda` and can be compared in the following histogram.
```{r, echo = FALSE, fig.width = 7, fig.asp = .62, fig.align = 'center'}
hist(exp.sims, breaks = 50, probability = F, 
     main="1000 simulations of exponential distribution means", 
     xlab="Spread")
abline(v = 1/lambda, col= 3, lwd = 0.75)
abline(v = exp.sims.mean, col = 2, lwd = 0.75)
legend("topright", 
       c(paste("Sample mean", 
               round(exp.sims.mean, 3)), 
         paste("Theoretical mean", 1/lambda)), 
       bty = "n",  
       lty = c(1,1), 
       col = c(col = 3, col = 2))
```
[See code here](#appendix-1)

## Distributions comparison
Compare how variable the sample is vs the expected distribution via variance

## Expected variance
[See code here](#appendix-2)

```{r, echo = FALSE}
matrix(c(1/lambda, 
         ((1/lambda)/sqrt(obs))^2, 
         exp.sims.mean, 
         var(exp.sims)), 
       nrow = 2,  
       dimnames = list(c("Mean", "Variance"), c("Expected", "Sample")))
```
## Showing distributions aproximately to the normal
```{r, echo = FALSE, fig.width = 7, fig.asp = .62, fig.align = 'center'}
set.seed(215)
hist(exp.sims, breaks = 50, probability = T,  
     main="1000 simulations of exponential distribution means", 
     xlab="Spread")
abline(v = 1/lambda,  col= 3, lwd = 0.75)
abline(v = exp.sims.mean, col = 2, lwd = 0.75)
legend("topright", 
       c(paste("Sample mean", round(exp.sims.mean, 3)),
         paste("Theoretical mean", 1/lambda)), 
       bty = "n",  
       lty = c(1,1),  
       col = c(col = 3, col = 2))
lines(density(exp.sims), col = 3)
lines(density(rnorm(1000, 5,((1/lambda)/sqrt(obs)))), col = 2)
```
[See code here](#appendix-3)

## Normal distribution aproximation
```{r,fig.width = 7, fig.asp = .62, fig.align = 'center'}
qqnorm(exp.sims)
qqline(exp.sims)
```


# Appendix

## Appendix 1
Code used in plot 1

<a id="Annex1"></a>
```{r, eval = FALSE, fig.width = 7, fig.asp = .62, fig.align = 'center'}
hist(exp.sims, breaks = 50, probability = F, 
     main="1000 simulations of exponential distribution means", 
     xlab="Spread")
abline(v = 1/lambda, col= 3, lwd = 0.75)
abline(v = exp.sims.mean, col = 2, lwd = 0.75)
legend("topright", 
       c(paste("Sample mean", 
               round(exp.sims.mean, 3)), 
         paste("Theoretical mean", 1/lambda)), 
       bty = "n",  
       lty = c(1,1), 
       col = c(col = 3, col = 2))
```


## Appendix 2
Code used in plot 2

```{r, eval = FALSE, fig.width = 7, fig.asp = .62, fig.align = 'center'}
set.seed(215)
hist(exp.sims, breaks = 50, probability = T,  
     main="1000 simulations of exponential distribution means", 
     xlab="Spread")
abline(v = 1/lambda,  col= 3, lwd = 0.75)
abline(v = exp.sims.mean, col = 2, lwd = 0.75)
legend("topright", 
       c(paste("Sample mean", round(exp.sims.mean, 3)),
         paste("Theoretical mean", 1/lambda)), 
       bty = "n",  
       lty = c(1,1),  
       col = c(col = 3, col = 2))
lines(density(exp.sims), col = 3)
lines(density(rnorm(1000, 5,((1/lambda)/sqrt(obs)))), col = 2)
```

## Appendix 3
Code used for the matrix

```{r, eval = FALSE}
matrix(c(1/lambda, 
         ((1/lambda)/sqrt(obs))^2, 
         exp.sims.mean, 
         var(exp.sims)), 
       nrow = 2,  
       dimnames = list(c("Mean", "Variance"), c("Expected", "Sample")))
```
