---
title: "Assignment Statistical Inference - Part 2"
author: "OscarAMtz"
date: "8 de abril de 2019"
output:
  pdf_document: default
  geometry: margin=2cm
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Part 2
##Exploratory data analyses in the native `ToothGrowth` database
The data set ToothGrowth contains measurements on the effects on tooth length (len) based on varying dosages of vitamin C for a cohort of guinea
pigs. [^1]

[^1]: Crampton, E. W. (1947). The growth of the odontoblast of the incisor teeth as a criterion of vitamin C intake of the guinea pig. The Journal of Nutrition, 33(5), 491–504. doi: 10.1093/jn/33.5.491 The Effect of Vitamin C on Tooth Growth in Guinea Pigs https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html

### Exploratory analysis
```{r}
data(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
unique(ToothGrowth$supp)
```

T test aplied to Tooth length ~ Supplement used, separed by dosage

```{r, echo = FALSE}
dose0.5 <- t.test(len ~ supp, ToothGrowth[(ToothGrowth$dose == 0.5),])
dose1 <- t.test(len ~ supp, ToothGrowth[(ToothGrowth$dose == 1),])
dose2 <- t.test(len ~ supp, ToothGrowth[(ToothGrowth$dose == 2),])

SummaryTStatistics <- data.frame(
  "p-value" = c(dose0.5$p.value, dose1$p.value, dose2$p.value),
  "CI95 low" = c(dose0.5$conf.int[1],dose1$conf.int[1], dose2$conf.int[1]),
  "CI95 high" = c(dose0.5$conf.int[2],dose1$conf.int[2], dose2$conf.int[2]),
  "Orange Juice mean" = c(dose0.5$estimate[[1]],dose1$estimate[[1]],dose2$estimate[[1]]),
  'Vitamin C mean' = c(dose0.5$estimate[[2]],dose1$estimate[[2]],dose2$estimate[[2]]),
  row.names = c("Dosage 0.5","Dosage 1","Dosage 2"))
SummaryTStatistics
```
[See code here](#appendix-1)


```{r, echo = FALSE, fig.width = 7, fig.asp = .62, fig.align = 'center'}
library(ggplot2)
ggplot(ToothGrowth, aes(factor(supp), len, fill = supp))+
  geom_boxplot() +
  facet_grid(.~dose) +
  xlab("Supplement") +
  ylab("Tooth length") +
  ggtitle("Supplement by tooth length per dosage") +
  theme(plot.title = element_text(hjust = 0.5)) 
```
[See code here](#appendix-2)

# Conclusion

As we can se **Orange Juice** it is more efective for tooth growth in lower dose than the **Vitamin C** with a significative p-value < 0.05 and Confident intervals over 0.

# Appendix

## Appendix 1
```{r, eval = FALSE}
dose0.5 <- t.test(len ~ supp, ToothGrowth[(ToothGrowth$dose == 0.5),])
dose1 <- t.test(len ~ supp, ToothGrowth[(ToothGrowth$dose == 1),])
dose2 <- t.test(len ~ supp, ToothGrowth[(ToothGrowth$dose == 2),])

SummaryTStatistics <- data.frame(
  "p-value" = c(dose0.5$p.value, dose1$p.value, dose2$p.value),
  "CI95 low" = c(dose0.5$conf.int[1],dose1$conf.int[1], dose2$conf.int[1]),
  "CI95 high" = c(dose0.5$conf.int[2],dose1$conf.int[2], dose2$conf.int[2]),
  "Orange Juice mean" = c(dose0.5$estimate[[1]],dose1$estimate[[1]],dose2$estimate[[1]]),
  'Vitamin C mean' = c(dose0.5$estimate[[2]],dose1$estimate[[2]],dose2$estimate[[2]]),
  row.names = c("Dosage 0.5","Dosage 1","Dosage 2"))
SummaryTStatistics
```

## Apendix 2
```{r, eval = FALSE, fig.width = 7, fig.asp = .62, fig.align = 'center'}
ggplot2::ggplot(ToothGrowth, aes(factor(supp), len, fill = supp))+
  geom_boxplot() +
  facet_grid(.~dose) +
  xlab("Supplement") +
  ylab("Tooth length") +
  ggtitle("Supplement by tooth length per dosage") +
  theme(plot.title = element_text(hjust = 0.5)) 
```
