---
title: "Activities Week 3"
author: "OscarAMtz"
date: "30 de enero de 2019"
output: html_document
---
## Debbuging R

- traceback : prints out the function call stack after an error occurs; does nothing if there´s no error

- debug : flags a function for "debug" mode which allows you tu step through execution of a function one line at a time

- browser :  suspends the execution of a function wherever it is called and puts the function in debug mode

- trace : allows you to insert debbuguing code into a function a specefic places

- recover : allows you to modify the error behavior so that you can browse the function call stack

### Diferent ways to loop functions in listed objects with the apply family

```{r}
with(mtcars, tapply(mpg, cyl, mean))
tapply(mtcars$mpg, mtcars$cyl, mean)
sapply(split(mtcars$mpg, mtcars$cyl), mean)

apply(iris[iris$Species == "virginica", 1:4], 2, mean)
```

## Example for scoping

```{r}
makeVector <- function(x = numeric()) {
        m <- NULL
        set <- function(y) {
                x <<- y
                m <<- NULL
        }
        get <- function() x
        setmean <- function(mean) m <<- mean
        getmean <- function() m
        list(set = set, get = get,
             setmean = setmean,
             getmean = getmean)
}

makeVector(9)

cachemean <- function(x, ...) {
        m <- x$getmean()
        if(!is.null(m)) {
                message("getting cached data")
                return(m)
        }
        data <- x$get()
        m <- mean(data, ...)
        x$setmean(m)
        m
}
x <- c(3,3.2,3.5,3.6)
vectormean <- makeVector()
```

## inverse of a matrix

```{r}
## Element for matrix structure
trn <- matrix(c(3,3.2,3.5,3.6), 2,2)
bw_trn_vec <- as.vector(trn)[length(trn):1]
matrix(bw_trn_vec, 2,2) 
### Matrix 2x2 example
x <- c(1:4)
dm <- sqrt(length(x))
mtx2x2 <- matrix(x,nrow = dm, ncol = dm)
matrix(as.vector(mtx2x2)[length(mtx2x2):1], dm, dm)
matrix(c(rep((c(1,-1), dm), rep(c(-1,1), dm))), nrow = dm, ncol = dm)
t(mtx2x2)

### Matrix 3X3 example
x <- c(1:9)
dm <- sqrt(length(x))
mtx3x3 <- matrix(x,nrow = dm, ncol = dm)
matrix(as.vector(mtx3x3)[length(mtx3x3):1], dm, dm)
matrix(rep(c(1,-1), dm^2), nrow = dm, ncol = dm)
t(mtx3x3)
### Matrix 4x4 example
x <- (1:16)
dm <- sqrt(length(x))
mtx4x4 <- matrix(x,nrow = dm, ncol = dm)
matrix(as.vector(mtx4x4)[length(mtx4x4):1], dm, dm)
t(mtx4x4)




x <- matrix(c(3,3.2,3.5,3.6),2,2)
matriz <- makeMatrix()
matriz$set(x)
cachematrix(matriz)
matriz$getmatrix(),
### Store the matrx
x <- c(3,3.2,3.5,3.6)
trn <- matrix(c(3,3.2,3.5,3.6), 2,2)
i_trn <- matrix(c(trn[2,2],trn[1,2]*-1,-1*trn[2,1],trn[1,1]),2,2)
inv_trn <- (1/(trn[1,1]*trn[2,2]-trn[1,2]*trn[2,1]))*i_trn
mat <- matrix(c(118.4, 135.2),2,1)
matrix(c(sum(inv_trn[1,1:2]*mat), sum(inv_trn[2,1:2]*mat)),2,1)
```

